### 2024-11-08 Analysis/AR1.Rmd
Addressed committee meeting comments - 08/11/2024

---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
library(tidyverse)
library(glmmTMB)
library(ggeffects)
```

# AR1

## Campelen

### Fall

```{r}
load(file = "./ModelSaves/FallCamp_Length.RData")
acc_surv <- read.csv("./Data/Capelin_AccSurv.csv")[1:2] %>% 
  filter(Year > 1994) %>% rename(acsurv_bm = X3L.Spring..Acoustic.Capelin.Biomass..kt.)

lr_dat <- data.frame(year = as.factor(acc_surv[2:28,1]), acc_bm = acc_surv[2:28,2], NLFPM_bm = rep$iye, dummy = as.factor(0))

fcar <- glmmTMB(scale(acc_bm) ~ NLFPM_bm + ar1(year+0|dummy), data = lr_dat)

summary(fcar)

plot(ggpredict(fcar, terms = c("NLFPM_bm")))+
  geom_point(data = lr_dat, aes(x = NLFPM_bm, y = acc_bm))

plot(lr_dat$acc_bm~lr_dat$NLFPM_bm)

```
Fall Campelen NLFPM correlation to acoustic - not significant positive relationship (Corr = 0.45, p = 0.0811)

### Spring

```{r}
load(file = "./ModelSaves/AP-SprCamp_Length.RData")
acc_surv <- read.csv("./Data/Capelin_AccSurv.csv")[1:2] %>% 
  filter(Year > 1994) %>% rename(acsurv_bm = X3L.Spring..Acoustic.Capelin.Biomass..kt.)

ap_lr_dat <- data.frame(year = as.factor(acc_surv[2:25,1]), acc_bm = acc_surv[2:25,2], NLFPM_bm = ap_rep$iye, dummy = as.factor(0))

# # Influential points
# ap_lr_dat <- data.frame(year = c(acc_surv[2:18,1], acc_surv[21:25,1]), acc_bm = c(acc_surv[2:18,2], acc_surv[21:25,2]), 
#                         NLFPM_bm = c(ap_rep$iye[2:18],ap_rep$iye[21:25]))

scapar <- glmmTMB(scale(acc_bm) ~ NLFPM_bm + ar1(year+0|dummy), data = ap_lr_dat)

summary(scapar)

plot(ggpredict(scapar, terms = c("NLFPM_bm")))+
  geom_point(data = ap_lr_dat, aes(x = NLFPM_bm, y = acc_bm))

plot(ap_lr_dat$acc_bm~ap_lr_dat$NLFPM_bm)

```
Spring Campelen AP NLFPM correlation to acoustic - not significant positive relationship (Corr = 0.54, p = 0.673)

```{r}
load(file = "./ModelSaves/GH-SprCamp_Length.RData")
acc_surv <- read.csv("./Data/Capelin_AccSurv.csv")[1:2] %>% 
  filter(Year > 1994) %>% rename(acsurv_bm = X3L.Spring..Acoustic.Capelin.Biomass..kt.)

gh_lr_dat <- data.frame(year = as.factor(acc_surv[2:25,1]), acc_bm = acc_surv[2:25,2], NLFPM_bm = gh_rep$iye, dummy = as.factor(0))

# gh_lr_dat <- data.frame(year = c(acc_surv[2:18,1], acc_surv[21:25,1]), acc_bm = c(acc_surv[2:18,2], acc_surv[21:25,2]), 
#                         NLFPM_bm = c(gh_rep$iye[2:18],gh_rep$iye[21:25]))

scghar <- glmmTMB(scale(acc_bm) ~ NLFPM_bm + ar1(year+0|dummy), data = gh_lr_dat)

summary(scghar)

plot(ggpredict(scghar, terms = c("NLFPM_bm")))+
  geom_point(data = gh_lr_dat, aes(x = NLFPM_bm, y = acc_bm))

plot(gh_lr_dat$acc_bm~gh_lr_dat$NLFPM_bm)

```
Spring Campelen GH NLFPM correlation to acoustic - No convergence

## Engel

### Fall
```{r}
load(file = "./ModelSaves/ACAP-FallEng_Length.RData")
acc_surv <- read.csv("./Data/Capelin_AccSurv.csv")[1:2] %>% 
  filter(Year > 1980 & Year < 1996) %>% rename(acsurv_bm = X3L.Spring..Acoustic.Capelin.Biomass..kt.)

# Set up relationship data frame - ACAP
ac_lr_dat <- data.frame(year = as.factor(acc_surv[2:15,1]), acc_bm = acc_surv[2:15,2], NLFPM_bm = ac_rep$iye, dummy = as.factor(0)) #have to y+1

feacar <- glmmTMB(scale(acc_bm) ~ NLFPM_bm + ar1(year+0|dummy), data = ac_lr_dat)

summary(feacar)

plot(ggpredict(feacar, terms = c("NLFPM_bm")))+
  geom_point(data = ac_lr_dat, aes(x = NLFPM_bm, y = acc_bm))

plot(ac_lr_dat$acc_bm~ac_lr_dat$NLFPM_bm)

```
Fall Engel ACAP NLFPM correlation to acoustic - significant positive relationship (Corr = 0.02, p = 0.0376)

```{r}
load(file = "./ModelSaves/GH-FallEng_Length.RData")
acc_surv <- read.csv("./Data/Capelin_AccSurv.csv")[1:2] %>% 
  filter(Year > 1980 & Year < 1996) %>% rename(acsurv_bm = X3L.Spring..Acoustic.Capelin.Biomass..kt.)

gh_lr_dat <- data.frame(year = as.factor(acc_surv[2:15,1]), acc_bm = acc_surv[2:15,2], NLFPM_bm = gh_rep$iye, dummy = as.factor(0))

feghar <- glmmTMB(scale(acc_bm) ~ NLFPM_bm + ar1(year+0|dummy), data = gh_lr_dat)

summary(feghar)

plot(ggpredict(feghar, terms = c("NLFPM_bm")))+
  geom_point(data = gh_lr_dat, aes(x = NLFPM_bm, y = acc_bm))

plot(gh_lr_dat$acc_bm~gh_lr_dat$NLFPM_bm)

```
Fall Engel GH NLFPM correlation to acoustic - No converge

### Spring

```{r}
load(file = "./ModelSaves/ACAP-SprEng_Length.RData")
acc_surv <- read.csv("./Data/Capelin_AccSurv.csv")[1:2] %>% 
  filter(Year > 1983 & Year < 1996) %>% rename(acsurv_bm = X3L.Spring..Acoustic.Capelin.Biomass..kt.)

ac_lr_dat <- data.frame(year = as.factor(acc_surv[2:13,1]), acc_bm = acc_surv[2:13,2], NLFPM_bm = c(ac_rep$iye[12], ac_rep$iye[1:11]), dummy = as.factor(0)) #have to y+1

seacar <- glmmTMB(scale(acc_bm) ~ NLFPM_bm + ar1(year+0|dummy), data = ac_lr_dat)

summary(seacar)

plot(ggpredict(seacar, terms = c("NLFPM_bm")))+
  geom_point(data = ac_lr_dat, aes(x = NLFPM_bm, y = acc_bm))

plot(ac_lr_dat$acc_bm~ac_lr_dat$NLFPM_bm)

```
Spring Engel ACAP NLFPM correlation to acoustic - no significant correlation (Corr = 0.03, p = 0.6177)

```{r}
load(file = "./ModelSaves/GH-SprEng_Length.RData")
acc_surv <- read.csv("./Data/Capelin_AccSurv.csv")[1:2] %>% 
  filter(Year > 1983 & Year < 1996) %>% rename(acsurv_bm = X3L.Spring..Acoustic.Capelin.Biomass..kt.)

gh_lr_dat <- data.frame(year = as.factor(acc_surv[2:13,1]), acc_bm = acc_surv[2:13,2], NLFPM_bm = c(gh_rep$iye[12], gh_rep$iye[1:11]), dummy = as.factor(0))

seghar <- glmmTMB(scale(acc_bm) ~ NLFPM_bm + ar1(year+0|dummy), data = gh_lr_dat)

summary(seghar)

plot(ggpredict(seghar, terms = c("NLFPM_bm")))+
  geom_point(data = gh_lr_dat, aes(x = NLFPM_bm, y = acc_bm))

plot(gh_lr_dat$acc_bm~gh_lr_dat$NLFPM_bm)

```
Spring Engel ACAP NLFPM correlation to acoustic - no significant correlation (Corr = 0.09, p = 0.907)